Prefetching SPARQL query cacher

### 2016-07-28 dissertation/cache-proxy.tex
start of an planning elaboration

\subsection{Key Problems}

>>>>

\subsection{Understanding the Query Planning}

We saw in Section~\ref{sec:introinteract} that a query would be parsed
to a tree of algebra objects and then a query planner would traverse
the algebra tree to create a plan tree by finding all the ways the
data can be accessed, or operations executed over them. For example,
if the input query has a BGP with two triple patterns and there are
two types of join operations, it may result in four trees, where each
of the join operation plans are paired with both access plans for the
triple patterns.

It is next the query planner's task to determine which of the possible
plans it should execute. A common choice to perform this task is to
estimate the cost executing various choices.

The system we propose is based on this paradigm, any choices we make
are therefore to be encoded into a cost model. The cost model itself
is not considered a part of the contributions, it is a simple cost
model based on heuristics that arises from the following
discussion. The heuristics themselves are detailed in
Section~\ref{sec:costheuristics}.

Recall Problem~\ref{prob:cachecartesian}, where a BGP with three
triple patterns could break into two components if the triple pattern
connecting the two others was cached. It is one of the requirements
for the heuristics of the cost model that it is be able to avoid such
situations.

As mentioned, the proposed query planner takes may access a remote
SPARQL endpoint and a Linked Data Fragments Server as well as the
local cache. I have also said that a key motivation behind the study
is to relieve the remote endpoint of load. It was argued in
\cite{verborgh2014querying} that LDF shifts the burden of query
evaluation to the client, and so relieves the server of load. It is
not clear from the literature that this strikes the right balance, it
is my belief that the balance must depend on the concrete server and
client capabilities, keeping in mind
Problem~\ref{prob:microcontroller}. Future work to determine the right
balance in a given situation would be welcome in this area, but I
believe it would be advantageous to build on a cost model.

Following this discussion, consider this example:
\begin{example}{A four-triple pattern BGP}
\begin{verbatim}
?s <http://example.org/m/p> "1" ;
   <http://example.org/m/q> <http://example.org/m/a> ;
   <http://example.org/m/p> ?o .
?o <http://example.org/m/b> "2" .
\end{verbatim}
\end{example}
Say that the first triple pattern has a result in the cache. This will
not cause Problem~\ref{prob:cachecartesian} to occur (it would have
occurred if the third triple pattern had a cached result, however), so
we can safely use the cache. The results of the three remaining triple
patterns must be fetched from a remote server. There are essentially
two reasonable choices: Either fetch the result from an LDF server as
argued by \cite{verborgh2014querying} and join them locally or send a
single, rewritten SPARQL query to the remote endpoint for the
evaluation. When the results are returned from both the cache and the
remote endpoint, they must then finally be joined to produce the final
result returned to the client. A string representation of the plan
tree from the current implementation reflects this and reads:
\begin{example}{An example Attean plan}
\begin{verbatim}
- Hash Join { s } (distinct cost: 74)
-   SPARQLBGP (cost: 60)
-     Quad { ?s, <http://example.org/m/p>, ?o, <http://test.invalid/graph> }
-     Quad { ?o, <http://example.org/m/b>, "2", <http://test.invalid/graph> }
-     Quad { ?s, <http://example.org/m/q>, <http://example.org/m/a>, <http://test.invalid/graph> }
-   Iterator (?s with 2 elements) (cost: 2)
\end{verbatim}
\end{example}
